Inference for Numerical Data

### 2018-10-30 DATA606/Inference for Numerical Data.Rmd
More work

@@

(a) Raina wants to use a 90% confidence interval. How large a sample should she collect?

(b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size, determine whether his sample should be larger or smaller than Raina’s, and explain your reasoning.

(c) Calculate the minimum required sample size for Luke.

@@

(a) Is there a clear difference in the average reading and writing scores?

(b) Are the reading and writing scores of each student independent of each other?

(c) Create hypotheses appropriate for the following research question: is there an evident difference in the average scores of students in the reading and writing exam?

(d) Check the conditions required to complete this test.

(e) The average observed difference in scores is $\bar{x}_{read−write}$ = −0.545, and the standard deviation of the differences is 8.887 points.  Do these data provide convincing evidence of a difference between the average scores on the two exams?

(f) What type of error might we have made? Explain what the error means in the context of the application.

@@

@@

@@

(a) Write hypotheses for evaluating whether the average number of hours worked varies across the five groups.

(b) Check conditions and describe any assumptions you must make to proceed with the test.

@@

|           |  Df | Sum Sq  | Mean Sq | F value | Pr(>F) |
|-----------|-----|---------|---------|---------|--------|
| degree    |** **| ** **   | 501.54  | ** **   | 0.0682 |
| Residuals |** **| 267,382 | ** **   |         |        |
| Total     |** **| ** **   |         |         |        |

(d) What is the conclusion of the test?

>>>>

```{r}
# Sample mean is midpoint of the confidence interval endpoints
x_bar <- mean(c(65,77))

# Margin of error is the difference between the midpoint and the CI endpoints
moe <- 77 - x_bar
```

$`r x_bar` \pm `r moe` = \bar{x} \pm Z \frac{s}{\sqrt{n}}$

Given z = 1.645 and n = 25

$`r moe` = 1.645 \frac{s}{\sqrt{25}} \rightarrow `r moe` = 1.645 \frac{s}{5} \rightarrow `r 5*moe` = 8.225 \times s \rightarrow s = \frac{`r 5*moe`}{8.225} \rightarrow s = `r (5*moe)/8.225`$

```{r}
# Given
z <- 1.645 # 90% confidence interval z-score
n <- 25

# Sample standard deviation (s)
s <- (sqrt(n) * moe) / (sqrt(n) * z)
```

**Sample mean: `r x_bar`;  Margin of Error: `r moe`; Sample Standard Deviation: `r s`;**

@@

(a) Raina wants to use a 90% confidence interval. How large a sample should she collect? 
```{r 90_ci_sample_size_Q5_14}
z <- 1.645 # 90% confidence interval z-score
sigma <- 250
moe <- 25

n <- ((z*sigma)/moe)^2
```

**Sample Size: `r n`**

$25 = 1.645 \frac{250}{\sqrt{n}} \rightarrow 25 = \frac{411.25}{\sqrt{n}} \rightarrow 1 = \frac{16.45}{\sqrt{n}} \rightarrow \sqrt{n} = 16.45 \rightarrow n = \pm 270.6025$

(b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size, determine whether his sample should be larger or smaller than Raina’s, and explain your reasoning. **Larger.  In order to have a higher confidence we would need a larger sample size.**

(c) Calculate the minimum required sample size for Luke.
collect?

```{r 99_ci_sample_size_Q5_14}
z <- 2.58 # 99% confidence interval z-score
sigma <- 250
moe <- 25

n <- ((z*sigma)/moe)^2
```

**Sample Size: `r n`**

@@

(a) Is there a clear difference in the average reading and writing scores? **Yes the reading score appears to be lower with more variation than the writing score.**

(b) Are the reading and writing scores of each student independent of each other? **No because the same student takes both tests.**

(c) Create hypotheses appropriate for the following research question: is there an evident difference in the average scores of students in the reading and writing exam? **$H_0 = \bar{x}_\text{read} - \bar{x}_\text{write} = 0$  and $H_A = \bar{x}_\text{read} - \bar{x}_\text{write} != 0$**

(d) Check the conditions required to complete this test. **Both distributions appear to be normally distributed and centered around zero.  The number of observations is larger than 30.**

(e) The average observed difference in scores is $\bar{x}_{read−write}$ = −0.545, and the standard deviation of the differences is 8.887 points.  Do these data provide convincing evidence of a difference between the average scores on the two exams? **The p-value suggests we should not reject the null hypothesis**

```{r}
x_bar <- -0.545
sigma <- 8.887
n <- 200
z <- (x_bar)/(sigma/sqrt(n))
1.96 * pnorm(-abs(z))
```

(f) What type of error might we have made? Explain what the error means in the context of the application.**A type II error.  We beleive there isn't a difference when there is.**

@@

**Yes.  Using an confidence level of 95%, the margin of error is:**

$MOE = Z\frac{s}{\sqrt{n}} \rightarrow MOE = 1.96\frac{8.887}{\sqrt{200}} \rightarrow MOE = 1.231675$

**Therefore the confidence interval is `r -0.545 - 1.231675` to `r -0.545 + 1.231675` which contains zero, so we cannot reject the null hypothesis.**

@@

```{r}
x_bar_automatic <- 16.12
s_automatic <- 3.58
n_automatic <- 26
x_bar_manual <- 19.85
s_manual <- 4.51
n_manual <- 26
point_estimate <- x_bar_automatic - x_bar_manual
df <- n_automatic - 1
se <- sqrt(((s_automatic^2)/n_automatic) + ((s_manual^2)/n_manual))
t_value <- point_estimate/se
critical_t <- pt(t_value, df)

c(point_estimate + (critical_t * 1.95), point_estimate - (critical_t * 1.95)) 
```

**The difference is roughly -3.727 and -3.733 mpg.**

@@

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/5_48.png)

(a) Write hypotheses for evaluating whether the average number of hours worked varies across the five groups. **H0: $\mu$1 = $\mu$2 = $\mu$3 = $\mu$4 = $\mu$5 and HA: The average varies across some or all groups.**

(b) Check conditions and describe any assumptions you must make to proceed with the test. **Independent, normaly distributed observations across groups.**

@@

|           |    Df   | Sum Sq  | Mean Sq | F value | Pr(>F) |
|-----------|---------|---------|---------|---------|--------|
| degree    |**     **| ** **   | 501.54  | ** **   | 0.0682 |
| Residuals |**     **| 267,382 | ** **   |         |        |
| Total     |**1,172**| ** **   |         |         |        |

(d) What is the conclusion of the test? **We cannot reject the null hypothesis.**
